Video is widely recorded but remains difficult to search at the moment
level, especially when a query expresses causal or temporal intent
rather than literal visual content. This dissertation presents Visual
RAG, an end-to-end agentic retrieval-augmented generation system that
indexes timestamped visual and transcript evidence, decomposes
natural-language questions into retrieval-friendly descriptions,
searches and re-ranks candidate moments, and answers with grounded
timestamp citations. The system is designed to run with modest resources
by separating an offline indexing stage from a lightweight online query
path.

The design is evaluated on 3,358 temporally grounded questions over 567
videos from NExT-QA and NExT-GQA. Visual retrieval substantially
outperforms transcript retrieval at moment granularity. Replacing the
baseline CLIP index with SigLIP and adding query decomposition nearly
doubles corpus-scope recall at rank one from 0.026 to 0.048. A two-stage
configuration using a cheaper index and a stronger visual re-ranker
recovers most of the quality of an expensive index. Two negative results
are also established: naive late fusion with sparse conversational
transcripts and text cross-encoder re-ranking both reduce retrieval
quality. At the answering stage, a bounded LangGraph agent improves
five-choice accuracy from 0.447 to 0.547, while a controlled experiment
on temporal questions shows that supplying visual evidence raises
accuracy from 0.341 to 0.636.

The results separate three bottlenecks in video RAG: recall, ranking
precision, and evidence delivery. Query decomposition, visual
re-ranking, and multimodal answering address these bottlenecks
independently and compose effectively. The dissertation contributes a
reproducible system, evaluation harness, ablation study, and practical
design guidance for temporally grounded video search on consumer
hardware.
